Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-06-22 survey/survey.tex
Next 2 sets of level-2 combinators. And also trailing blanks.

\begin{figure}[t]
\[
\Rule{}
{\jdg{}{}{c_1 : t_1 \iso t_2}}
{\jdg{}{}{\ !\ c_1 : t_2 \iso t_1}}
{}
\]
\caption{Derived $\Pi$-combinator.}
\label{derived-pi-combinators}
\end{figure}

@@

@@

\begin{figure}[tb]
Let $c_0 : 0 \leftrightarrow 0$, $c_1 : 1 \leftrightarrow 1$, and $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlp \odot c \Leftrightarrow (c_0 \oplus c) \odot \identlp}
\qquad
  {\identrp \odot (c_0 \oplus c) \Leftrightarrow c \odot \identrp}
\\
  {\identlsp \odot c \Leftrightarrow (c \oplus c_0) \odot \identlsp}
\qquad
  {\identrsp \odot (c \oplus c_0) \Leftrightarrow c \odot \identrsp}
\\
  {\identlt \odot c \Leftrightarrow (c_1 \otimes c) \odot \identlt}
\qquad
  {\identrt \odot (c_1 \otimes c) \Leftrightarrow c \odot \identrp}
\\
  {\identlst \odot c \Leftrightarrow (c \otimes c_1) \odot \identlst}
\qquad
  {\identrst \odot (c \otimes c_1) \Leftrightarrow c \odot \identrst}
\\
  {\identlt \Leftrightarrow \distl \odot (\identlt \oplus \identlt)}
\\
\identlp \Leftrightarrow \swapp \odot \identrsp
\qquad
\identlt \Leftrightarrow \swapt \odot \identrst
\end{array}\]
\caption{\label{figg}Signatures of level-2 $\Pi$-combinators: unit}
\end{figure}

The constructs in Fig.~\ref{figg} \ldots 
\begin{center}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.8}]
\end{tikzpicture}
\end{center}

The constructs in Fig.~\ref{figf} \ldots 
\begin{center}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.8}]
\end{tikzpicture}
\end{center}

>>>>

\begin{figure}[t]
\[
\Rule{}
{\jdg{}{}{c_1 : t_1 \iso t_2}}
{\jdg{}{}{\ !\ c_1 : t_2 \iso t_1}}
{}
\]
\caption{Derived $\Pi$-combinator.}
\label{derived-pi-combinator}
\end{figure}

@@

\begin{figure}[tb]
Let $c_0 : 0 \leftrightarrow 0$, $c_1 : 1 \leftrightarrow 1$, and $c : t_1 \leftrightarrow t_2$:
\[\def\arraystretch{1.3}
\begin{array}{c}
  {\identlp \odot c \Leftrightarrow (c_0 \oplus c) \odot \identlp}
\qquad
  {\identrp \odot (c_0 \oplus c) \Leftrightarrow c \odot \identrp}
\\
  {\identlsp \odot c \Leftrightarrow (c \oplus c_0) \odot \identlsp}
\qquad
  {\identrsp \odot (c \oplus c_0) \Leftrightarrow c \odot \identrsp}
\\
  {\identlt \odot c \Leftrightarrow (c_1 \otimes c) \odot \identlt}
\qquad
  {\identrt \odot (c_1 \otimes c) \Leftrightarrow c \odot \identrp}
\\
  {\identlst \odot c \Leftrightarrow (c \otimes c_1) \odot \identlst}
\qquad
  {\identrst \odot (c \otimes c_1) \Leftrightarrow c \odot \identrst}
\\
  {\identlt \Leftrightarrow \distl \odot (\identlt \oplus \identlt)}
\\
\identlp \Leftrightarrow \swapp \odot \identlsp
\qquad
\identlt \Leftrightarrow \swapt \odot \identlst
\end{array}\]
\caption{\label{figg}Signatures of level-2 $\Pi$-combinators: unit}
\end{figure}

The constructs in Fig.~\ref{figg} may at first blush look similarly straightforward,
but deserve some pause. One obvious question: What is the point of
$c_0 : 0 \leftrightarrow 0$, isn't that just the identity combinator $\idc$
for $A = 0$ (as defined in Fig.~\ref{type-isos})? Operationally, $c_0$
is indeed indistinguishable from $\idc$. However, there are multiple syntactic
ways of writing down combinators of type $0 \leftrightarrow 0$, and the
first combinator in Fig.~\ref{figg} applies to all of them uniformly.
This is another subtle aspect of coherence: all reasoning must be valid for
all possible models, not just the one we have in mind. So even though
operational reasoning may suggest that some relations \emph{may} be
true between combinators, it can also mislead. The same reasoning
applies to $c_1 : 1 \leftrightarrow 1$.  The first $8$ combinators can
then be read as basic coherence for unit introduction and elimination,
in both additive and multiplicative cases.

The last two capture
another simple idea, related to swapping: eliminating a unit
on the left is the same as first swapping then eliminating on the
right (both additively and multiplicatively). As a side note,
these are not related to \emph{commutativity}, but rather
come from one of the simplest coherence condition for
braided monoidal categories. In other words, it reflects the
idempotence of $\swapp$ and $\swapt$ rather than the
commutativity of $\oplus$ and $\otimes$.

The first two equivalences in Fig.~\ref{figf} reflect the basic
coherence between type-level swapping and and the level-2 combinator
actions. The next four arise because of interactions between
(additive and multiplicative) level-1 associativity and swapping.
In other words, they arise as critical pairs.  For example,
the first expresses that the two ways of going from
$\left(A \oplus B\right) \oplus C$ to $B \oplus \left(C \oplus A\right)$
are equivalent, with the second saying that the reverse (i.e.
the results of applying $!$\~) also gives equivalent programs.
The last two say the same but for the multiplicative structure.

@@
